MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites \\ Supplementary Material

### 2014-12-08 chapters/MetaTracts.tex
shortened introduction

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel. But the application areas of CFRPs are not limited on these two industries. For a wide range of different areas ranging from marine and civil engineering, wind energy, to sporting goods manufacturing, CFRP is highly interesting. The main reason therefore is found in CFRP materials having desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. CFRP materials show these characteristics at considerably lower weight compared to conventional materials. At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs. Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates addressed in this work, consist of two main components:
a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. After curing, the production process of the CFRP laminate is finished.

The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}. 

@@

 We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. For example, spatial context answers questions such as: Are these bundles in contact at a particular position in the data? What is the relative orientations of the contacting bundles?

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data.
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, then we cluster the MetaTracts to create fiber bundles and finally generate visualizations from the clusters. For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

>>>>

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of these continuously rising demands. 
%CFRPs have been successfully introduced in aeronautic and automotive %applications within recent years, replacing parts made of conventional %materials such as aluminum or steel. But the application areas of CFRPs %are not limited on these two industries. For a wide range of different %areas ranging from marine and civil engineering, wind energy, to sporting %goods manufacturing, CFRP is highly interesting. 
%The main reason therefore is found in CFRP materials having desirable %characteristics such as high specific stiffness, high specific strength %and high corrosion resistance. 
CFRP materials are increasingly replacing conventional materials such as aluminum or steel, as they have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. 
Moreover, CFRP materials show these characteristics at considerably lower weight. At the same time, highly complex and integrated components, which were previously impossible to manufacture may be produced from CFRPs. Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components and more specifically CFRP laminates with endless carbon fibers as addressed in this work, consist of two main components:
a matrix, which acts as a bonding component, and the reinforcements, which allow for achieving the desired characteristics. Various production processes are used to manufacture CFRP laminates. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth are also referred to as fabric and form the geometrical structure of the final CFRP materials. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth in similar or different orientation. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of the matrix is to act as a bonding between the individual carbon fiber bundles. After curing, the production process of the CFRP laminate is finished.

The increasing share of CFRPs as well as the complexity of material system and final components has generated a strong demand towards non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing (UT) is the most commonly used method for this purpose. While UT provides a quick and cost-efficient overview of the material, it also shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. In contrast to UT, XCT generates a highly detailed 3D volumetric representation of the scanned specimen. In cone beam XCT geometry the specimen is placed on a rotary table between X-ray source and detector. The X-rays passing through the specimen get attenuated by the materials present. By transferring the X-rays in a scintillator layer into visible light, the detector records the corresponding 2D attenuation image (penetration image). The specimen is rotated stepwise recording at each position a 2D attenuation image to generate the full series of attenuation images of a $360^\circ$ rotation for a complete reconstruction of the data volume ~\cite{heinzl-2008-thesis}. Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen. 
While cone beam XCT can reach voxel sizes below 500 $nm$ generating high resolution volume data for comprehensive and detailed analyses, unfortunately there is still a trade-off between viewport and image resolution. The magnification reached within an XCT scan is determined by the used distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: while higher resolutions decrease the viewport but show more details, lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

Figure~\ref{fig:data-char} provides an example of our targeted dataset type. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: Firstly, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. Secondly, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}. 

@@

We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. It provides answers to questions such as: Are these bundles in contact at a particular position in the data? What is the relative orientations of the contacting bundles?

Providing answers to the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We present and evaluate a method which uses visualization techniques to gain insight into our data.
We interpret and advance techniques from diffusion tensor imaging to extract and visualize geometric structures from XCT data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts. Then we cluster the MetaTracts to generate and visualize fiber bundles. For a detailed description of the MetaTracts generation process see Sec.~\ref{sec:approach}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
